Toward a secure TCP/IP stack --- Technical report

### 2020-07-27 report/report.tex
Update report according to Yannick's comments

        \draw ($(listen.south) + (-1em,0)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=1em]{SYN/SYN + ACK} 
        ($(synRcvd.east) + (0,1em)$);
        \draw ($(listen.south) + (1em,0)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1em]{\emph{Send}/SYN} 
        ($(synSent.west) + (0,1em)$);

@@

        \draw ($(synRcvd.east) + (0,-1em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=-1em]{ACK/-} 
        ($(established.north) + (-1em,0)$);
        \draw ($(synSent.west) + (0,-1em)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{SYN + ACK/ACK} 
        ($(established.north) + (1em,0)$);

@@

        \draw (finWait1.south) 
        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK/-} 
        (finWait2.north);
        \draw ($(finWait1.east) + (0,-.5em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=1em]{FIN/ACK} 
        (closing.north);
        \draw (finWait1.south east) 
        edge[stateEdge] node[edgeLabel, xshift=1em, yshift=2em, text width=3em]{FIN + ACK/ACK} 
        (timeWait.north west);

@@

    We present a scenario on how a connection can be established between two TCPs,
    when one TCP (TCP A) is in the state \CLOSED{} and the other (TCP B) is in the state \LISTEN.
    \begin{enumerate}
        \item TCP A wants to initiate a connection with TCP B. It sends a SYN segment to
            TCP B and passes to the state \SYNSENT{}.

        \item TCP B receives a segment SYN. It has to respond to this segment by a segment
            containing the flags SYN and ACK (acknowledgment of the received segment).
            It changes its state for \SYNRECEIVED.

        \item When TCP A receives the segment sent by TCP B containing the flags SYN and ACK,
            it only has to send back an ACK of this segment to pass in the state \ESTABLISHED.

        \item TCP B receives the ACK of its segment and pass in state \ESTABLISHED{} too.
    \end{enumerate}
    At the end of this procedure, the two TCPs are in state \ESTABLISHED{} and can start to
    send data. Figure \ref{fig:handshake} illustrates the procedure described above. This procedure to open a connection is known as the ``three-way
    handshake'' in the TCP norm.

    \begin{figure}
        \centering
        \begin{tikzpicture}
            \node (A) at (0, 5) {\bfseries TCP A};
            \node (B) at (5, 5) {\bfseries TCP B};
            \draw[dashed] (-.7,4.8) -- (5.7,4.8);
            \node (TCPA1) at (0, 4.4) {\CLOSED{}};
            \node (TCPB1) at (5, 3.9) {\LISTEN{}};
            \node (TCPA2) at (0, 3.4) {\SYNSENT};
            \node (TCPB2) at (5, 2.9) {\SYNRECEIVED{}};
            \node (TCPA3) at (0, 2.4) {\ESTABLISHED{}};
            \node (TCPB3) at (5, 1.9) {\ESTABLISHED{}};

            \draw[->] (TCPA1) -- (TCPA2);
            \draw[->] (TCPB1) -- (TCPB2);
            \draw[->] (TCPA2) -- (TCPA3);
            \draw[->] (TCPB2) -- (TCPB3);
            \draw[->, dashed] (0.2, 3.9) -- node [midway, fill=white] {\footnotesize SYN} (TCPB1);
            \draw[->, dashed] (4.8, 3.4) -- node [midway, fill=white] {\footnotesize SYN + ACK} (TCPA2);
            \draw[->, dashed] (0.2, 2.9) -- node [midway, fill=white] {\footnotesize ACK} (TCPB2);
        \end{tikzpicture}
        \par
        \caption{Three-way hanshake procedure. The solid lines represent the transition between
        the states and the dashed lines represent the messages sent.}
        \label{fig:handshake}
    \end{figure}

@@

    This incorrect behavior has been detected thanks to the use of Klee and the improvement
    on how function's closure are computed. The contract for the call
    \begin{lstlisting}[language=Ada, basicstyle=\ttfamily\small]
    Tcp_Wait_For_Events (Sock, SOCKET_EVENT_TX_READY,
                         Sock.S_Timeout, Event);
    \end{lstlisting}
    was not correct at first, because it was written by hands and thus not enough precise.
    Then, the possible state at the end of the function \lstinline[language=Ada]{Tcp_Change_State}
    was reduce to \ESTABLISHED{} and the same problem existed for the call
    at line \ref{bugProg:tcpWaitForEvents}. At this point no problem was detected with SPARK,
    because the contracts were true. With Klee, some transitions that were not considered at first,
    because hide in subfunctions have been discovered. In particular there exists a transition from
    \SYNRECEIVED{} to \CLOSEWAIT{} if a FIN flags is present in the same segment of the ACK that was not
    spotted before. Putting these information in the function \lstinline[language=Ada]{Tcp_Process_One_Segment},
    and proving the post-condition of the function \lstinline[language=Ada]{Tcp_Wait_For_Events}
    directly with SPARK using algorithm \ref{algo:waitForEvents}, rather than not proving it like
    it was done before, helps to correct the contracts and find the bug.

@@

    Other improvements that could be done in the future (this is a not exhaustive aggregate of ideas,
    this list is subject to change):
    \begin{itemize}
        \item Fully translate the file \texttt{tcp\_misc.c} in Ada to have more guarantees.
        \item Same for the file \texttt{tcp\_fsm.c} that has been investigated with klee. But
        it is not enough and we need more guarantees for this file, in order to prove the
        interactions between tasks first, and also to prove the correctness of this part
        of code w.r.t. the TCP norm.
        \item The translation only considers one possible preprocessing. A future work would
        be considering all the possible preprocessing by adding constants like in the C
        version. An harder work is to find a way to keep a coherence between the C preprocessing
        and the Ada preprocessing. It can be achieved with an automatic tool, for example.

>>>>

        \draw ($(listen.south) + (-1em,0)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=1em]{SYN/(SYN \& ACK)} 
        ($(synRcvd.east) + (0,1em)$);
        \draw ($(listen.south) + (1em,0)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1em]{\emph{Send}/SYN} 
        ($(synSent.west) + (0,1em)$);

@@

        \draw ($(synRcvd.east) + (0,-1em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=-1em]{ACK/-} 
        ($(established.north) + (-1em,0)$);
        \draw ($(synSent.west) + (0,-1em)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{(SYN \& ACK)/ACK} 
        ($(established.north) + (1em,0)$);

@@

        \draw (finWait1.south) 
        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK/-} 
        (finWait2.north);
        \draw ($(finWait1.east) + (0,-.5em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=1em]{FIN/ACK} 
        (closing.north);
        \draw (finWait1.south east) 
        edge[stateEdge] node[edgeLabel, xshift=1em, yshift=2em, text width=3em]{(FIN \& ACK)/ACK} 
        (timeWait.north west);

@@

    We present a scenario on how a connection can be established between two TCPs,
    when one TCP (TCP $\alpha$) is in the state \CLOSED{} and the other (TCP $\beta$) is in the state \LISTEN.
    \begin{enumerate}
        \item TCP $\alpha$ wants to initiate a connection with TCP $\beta$. It sends a SYN segment to
            TCP $\beta$ and passes to the state \SYNSENT{}.

        \item TCP $\beta$ receives a segment SYN. It has to respond to this segment by a segment
            containing the flags SYN and ACK (acknowledgment of the received segment).
            It changes its state for \SYNRECEIVED.

        \item When TCP $\alpha$ receives the segment sent by TCP $\beta$ containing the flags SYN and ACK,
            it only has to send back an ACK of this segment to pass in the state \ESTABLISHED.

        \item TCP $\beta$ receives the ACK of its segment and pass in state \ESTABLISHED{} too.
    \end{enumerate}
    At the end of this procedure, the two TCPs are in state \ESTABLISHED{} and can start to
    send data. Figure \ref{fig:handshake} illustrates the procedure described above.
    This procedure to open a connection is known as the ``three-way handshake'' in the TCP norm.

    \begin{figure}
        \centering
        \begin{tikzpicture}
            \node (A) at (0, 5) {\bfseries TCP $\alpha$};
            \node (B) at (5, 5) {\bfseries TCP $\beta$};
            \draw[dashed] (-.7,4.8) -- (5.7,4.8);
            \node (TCPA1) at (0, 4.4) {\CLOSED{}};
            \node (TCPB1) at (5, 3.9) {\LISTEN{}};
            \node (TCPA2) at (0, 3.4) {\SYNSENT};
            \node (TCPB2) at (5, 2.9) {\SYNRECEIVED{}};
            \node (TCPA3) at (0, 2.4) {\ESTABLISHED{}};
            \node (TCPB3) at (5, 1.9) {\ESTABLISHED{}};

            \draw[->] (TCPA1) -- (TCPA2);
            \draw[->] (TCPB1) -- (TCPB2);
            \draw[->] (TCPA2) -- (TCPA3);
            \draw[->] (TCPB2) -- (TCPB3);
            \draw[->, dashed] (0.2, 3.9) -- node [midway, fill=white] {\footnotesize SYN} (TCPB1);
            \draw[->, dashed] (4.8, 3.4) -- node [midway, fill=white] {\footnotesize SYN + ACK} (TCPA2);
            \draw[->, dashed] (0.2, 2.9) -- node [midway, fill=white] {\footnotesize ACK} (TCPB2);
        \end{tikzpicture}
        \par
        \caption{Three-way hanshake procedure. The solid lines represent the transitions between
        the states and the dashed lines represent the messages sent.}
        \label{fig:handshake}
    \end{figure}

@@

    This incorrect behavior has been detected thanks to the use of KLEE and the improvement
    on how function's closure are computed. The contract for the call
    \begin{lstlisting}[language=Ada, basicstyle=\ttfamily\small]
    Tcp_Wait_For_Events (Sock, SOCKET_EVENT_TX_READY,
                         Sock.S_Timeout, Event);
    \end{lstlisting}
    was not correct at first, because it was written by hands and thus not enough precise.
    Then, the possible state at the end of the function \lstinline[language=Ada]{Tcp_Change_State}
    was reduce to \ESTABLISHED{} and the same problem existed for the call
    at line \ref{bugProg:tcpWaitForEvents}. At this point no problem was detected with SPARK,
    because the contracts were true. With KLEE, some transitions that were not considered at first,
    because hide in subfunctions have been discovered. In particular there exists a transition from
    \SYNRECEIVED{} to \CLOSEWAIT{} if a FIN flags is present in the same segment of the ACK that was not
    spotted before. Putting these information in the function \lstinline[language=Ada]{Tcp_Process_One_Segment},
    and proving the post-condition of the function \lstinline[language=Ada]{Tcp_Wait_For_Events}
    directly with SPARK using algorithm \ref{algo:waitForEvents}, rather than not proving it like
    it was done before, helps to correct the contracts and find the bug.

@@

    Other improvements that could be done in the future (this is a not exhaustive aggregate of ideas,
    this list is subject to change):
    \begin{itemize}
        \item Fully translate the file \texttt{tcp\_misc.c} in Ada to have more guarantees.
        \item Same for the file \texttt{tcp\_fsm.c} that has been investigated with KLEE. But
        it is not enough and we need more guarantees for this file, in order to prove the
        interactions between tasks first, and also to prove the correctness of this part
        of code w.r.t. the TCP norm.
        \item The translation only considers one possible preprocessing. A future work would
        be considering all the possible preprocessing by adding constants like in the C
        version. An harder work is to find a way to keep a coherence between the C preprocessing
        and the Ada preprocessing. It can be achieved with an automatic tool, for example.
